---
title: "Report Example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Report Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

dir <- file.path(tempdir(), "example-report")
dir.create(dir)
knitr::opts_knit$set(root.dir = dir)

library(usethis)
library(RSQLite)
library(DBI)
```

This vignette will run through the steps needed to produce an Excel report from 
SQL using {quickReport}. We'll be looking at folder structures a bit, so we'll
load {fs} too:
```{r load_pkgs}
library(quickReport)
library(fs)
```

# Setting up a report project
Now we'll use `create_report()` to set up a project with the recommended folder
structure:
```{r create_proj, results='hide', message=FALSE}
create_report(path = ".")
```

`create_report()` creates a project with several non-standard files and folders:

- `Excel` is used for storing Excel templates (and comes with one you can edit)
- `Outputs` is the default folder where {quickReport} will try to place created
- `SQL` is used for storing SQL scripts that will be used to create a report
- `config.yml` can be used to configure an SQL database (powered by the
[{config}](https://rstudio.github.io/config/) package)
- `run-report.R` is an R script containing the minimal code needed to run the 
report.

```{r inspect_dir_1}
dir_tree()
```

# Setting up SQL
{quickReport} provides a function to create an example SQLite database in memory
containing 3 base datasets: `mtcars`, `beaver1` and `C02`.
```{r setup_sql_db}
db <- example_db()
RSQLite::dbListTables(db)
```

Let's add some SQL scripts to the SQL directory. These are the files that will 
be used to create a report when we run `report_from_sql()`
```{r make_sql_scripts}
# Define some SQL queries
queries <- list(
  "Active Beavers.sql" = "select * from beaver1 where activ = 1;",
  "High MPG Cars.sql"  = "select mpg, cyl, disp from mtcars where mpg >= 30;",
  "Mean C02 Uptake.sql" = "
    select Treatment, avg(uptake) as uptake
    from CO2
    group by Treatment;
  "
)

# Write the queries to the SQL directory
for (q in names(queries)) {
  readr::write_file(queries[[q]], file.path("SQL", q))
}
```

Let's check the SQL directory to confirm the files were written correctly:
```{r view_sql_dir}
dir_tree("SQL")
```

# Running the Report
We now have what we need to create a {quickReport} report from SQL: a
project with the correct file structure and an SQL database. With these in
place, we can create a fairly decent report by simply running
`report_from_sql()`:
```{r run_report_1}
report_1 <- nice_filename("demo-report", dir = "Outputs")
report_from_sql(report_1, sql_connection = db, file_template = NULL)
```

We can see that a file ``r report_1`` has been created. Let's check the report
to see what's been written on each sheet. Notice that the sheets have been 
given the same names as our SQL scripts:
```{r inspect_report_1}
read_worksheets(report_1)[-1]
```

# Using Transformers
{quickReport} allows the use of 'transformer' functions when producing a report.
These can be very useful when reporting from SQL, as many manipulations are much
easier to do using R code instead of SQL. Let's rerun the report, but this time 
we'll apply some transformers too:
```{r run_report_2, R.options=list(qr.quiet = TRUE)}
library(dplyr, warn.conflicts = FALSE)

# 'High MPG Cars' gets its own transformer that adds an extra column, while 
# the unnamed function will be applied to all other datasets
transformers <- list(
  "High MPG Cars" = function(data) data %>% mutate(best_mpg = mpg == max(mpg)),
  function(data) data %>% mutate(across(where(is.numeric), round))
)

# Generate a second report using transformations
report_2 <- nice_filename("demo-report-transformed", dir = "Outputs")
report_from_sql(
  report_2, sql_connection = db, transformers = transformers, 
  file_template = NULL
)
```

Looking at this second report, (conveniently named ``r report_2``), we can 
confirm that the transformations have been applied:
```{r inspect_report_2}
read_worksheets(report_2)[-1]
```

To finish, let's inspect the directory once more to see how our report looks:
```{r inspect_dir_2}
dir_tree()
```

